---
title: "rmdfileproject copie"
author: "victor souob"
date: "16/08/2020"
output: word_document
---
---
title: "EdxMovielensProject"
author: "victor souob"
date: "15/08/2020"
output: word_document
---

EDX MOVIE PROJECT : MOVIELENSE

In this project, the aim is to create a model of movie rating with the movielense data provided. The challenge is to create the model with a RMSE < 0.86490.

First, Let's download all the datasets, Libraries and packages. 
Then we will create a data partiton of the movilense ratings, wich 90 % of that partition will be  the training set (edx) and 10 % of that partition will be the test set (Validation).

```{r}

if(!require(tidyverse)) install.packages("tidyverse", repos = "http://cran.us.r-project.org")
if(!require(caret)) install.packages("caret", repos = "http://cran.us.r-project.org")
if(!require(data.table)) install.packages("data.table", repos = "http://cran.us.r-project.org")

library(tidyverse)
library(caret)
library(data.table)

install.packages("kableExtra")
library(kableExtra)


# MovieLens 10M dataset:
# https://grouplens.org/datasets/movielens/10m/
# http://files.grouplens.org/datasets/movielens/ml-10m.zip

dl <- tempfile()
download.file("http://files.grouplens.org/datasets/movielens/ml-10m.zip", dl)

ratings <- fread(text = gsub("::", "\t", readLines(unzip(dl, "ml-10M100K/ratings.dat"))),
                 col.names = c("userId", "movieId", "rating", "timestamp"))

movies <- str_split_fixed(readLines(unzip(dl, "ml-10M100K/movies.dat")), "\\::", 3)
colnames(movies) <- c("movieId", "title", "genres")

# if using R 3.6 or earlier
movies <- as.data.frame(movies) %>% mutate(movieId = as.numeric(levels(movieId))[movieId],
                                           title = as.character(title),
                                           genres = as.character(genres))
# if using R 4.0 or later
movies <- as.data.frame(movies) %>% mutate(movieId = as.numeric(movieId),
                                           title = as.character(title),
                                           genres = as.character(genres))

movielens <- left_join(ratings, movies, by = "movieId")

### DATA PARTITION 

# Validation set will be 10% of MovieLens data
set.seed(1, sample.kind="Rounding")

# if using R 3.5 or earlier, use `set.seed(1)` instead

test_index <- createDataPartition(y = movielens$rating, times = 1, p = 0.1, list = FALSE)
edx <- movielens[-test_index,]
temp <- movielens[test_index,]

# Make sure userId and movieId in validation set are also in edx set

validation <- temp %>% 
  semi_join(edx, by = "movieId") %>%
  semi_join(edx, by = "userId")

# Add rows removed from validation set back into edx set

removed <- anti_join(temp, validation)
edx <- rbind(edx, removed)

# Clean up memory by deleting unsused objects and performing a garbage collection
rm(dl, ratings, movies, test_index, temp, movielens, removed)
gc()

```

DATA ANALYSIS AND EXPLORATION

let's have a quick overview of the data , here we looking at all the variables to identify key variable for our model prediction.


```{r}
dim(edx)

head(edx)

```

Now Let's count all the movies in the data set and have an overview and tendencies.
```{r}
# let's group all movies by movieID 

edx_movies <- edx %>%
  group_by(movieId) %>%
  summarize(count = n()) %>%
  arrange(desc(count))

# let's have an overview distribution of movies in the data set

summary(edx_movies$count)

```
Here we can see that one movie was rated 31362 times. 
Also 122 movies represent half of the ratings
Movies was rated 842.9 in averages

NOW let's visualize the ratings distribution in the data set and also the movie rating distribution per year.
```{r}
# ratings Distribution

ggplot(data = edx, aes(x = rating)) +
  geom_bar() + 
  labs(title = "Ratings Distribution", x = "Rating", y = "Number of ratings")

# Movie Ratings distribution  per year 

movies_year <- edx %>%
  transform(timestamp = format(as.POSIXlt(timestamp, origin = "1970-01-01"), "%Y")) %>%
  select(timestamp, movieId) %>%
  group_by(timestamp) %>%
  summarise(count = n_distinct(movieId))

ggplot(data = movies_year, aes(x = timestamp, y = count)) +
  geom_bar(stat = "identity") + 
  labs(title = "Movies Ratings distribution per year", x = "Year", y = "Number of ratings")

```
-FINDING THE MODEL 
We are starting with a model , assuming  that all movies in the trainig set have equal ratings.
Then the formula for that model will be : Ymu,i= u + ϵ u,i
Here u represent the average rating for all movies and users in edx , and  ϵ represent all errors (in this model we are minimizing ϵ).
Now we can compute the average ratings on edx (u), test it into the validation set and predict the RMSE.
```{r}
 # we calculate the overall average rating on the training dataset
u <- mean(edx$rating)


# Here is the formula of RMSE
RMSE <- function(true_ratings = NULL, predicted_ratings = NULL) {
  sqrt(mean((true_ratings - predicted_ratings)^2))
}


# Calculate RMSE using validation ratings
  
  RMSE(validation$rating, u) # 1.06 as RMSE is not good enough
  
  
```
This model give us a RMSE of 1.06, which is not the goal for our ideal model (RMSE < 0.86490.).

MOVIE EFFECT
WE can Optimize our model by including the movie effect.
lets call bi the average rating of movie i.
To calculate the new model we can use the formula : 
Yu, i = u + bi + ϵui
if rearange the formula and isolate bi we will have :
bi = Yu,i - u. This means we can calculate bi by substracting the overall average of each movie rating with the overall average rating of all movies.
```{r}
# caluclate b_i for each movie and let's compare it  with the overall average u on training dataset

b_i <- edx %>%
  group_by(movieId) %>%
  summarize(b_i = mean(rating - u))


#  Lets add b_i into the validaation set and lets predict all unknown ratings with u and b_i

predicted_ratings <- validation %>% 
  left_join(b_i, by='movieId') %>%
  mutate(pred = u + b_i) %>%
  pull(pred)


#  calculate RMSE of movie ranking effect
RMSE(validation$rating, predicted_ratings) # 0.94 still not good enough

# plot the distribution of b_i's
qplot(b_i, data = b_i, bins = 15, color = I("black"))

```

MOVIE EFFECT AND USER EFFECT

We can include the user effect (bu) into the model to optimize it.
Yu, i = μ + bi + bu + ϵμ, i
we can then compute bu based on the formula above

```{r}

# lets train the model with movie effect (b_i) and users effect (b_u)

# let's find b_u

b_u <- edx %>% 
  left_join(b_i, by='movieId') %>%
  group_by(userId) %>%
  summarize(b_u = mean(rating - u - b_i))


# predict new ratings with movie and user bias

predicted_ratings <- validation %>% 
  left_join(b_i, by='movieId') %>%
  left_join(b_u, by='userId') %>%
  mutate(pred = u + b_i + b_u) %>%
  pull(pred)


# calculate RMSE of movie ranking effect
RMSE(predicted_ratings, validation$rating) # 0.8653488 we getting close 




```

LETS TRAIN THE MODEL WITH THE BEST REGULARIZATION FACTOR LAMBDA

```{r}
# lets optimized movie and user effect method with the best regularization factor (lamba)
  
  # let's determine the best lambda from a sequence
  
  lambdas <- seq(from=0, to=10, by=0.25 )
  
  # output RMSE of each lambda, repeat earlier steps (with regularization)
  
  rmses <- sapply (lambdas, function(l) {
    
    # calculate average rating across training data
    u <- mean(edx$rating)
    
    # compute regularized movie bias term
    b_i <- edx %>% 
      group_by(movieId) %>%
      summarize(b_i = sum(rating - u)/(n()+l))
    
    # compute regularize user bias term
    b_u <- edx %>% 
      left_join(b_i, by="movieId") %>%
      group_by(userId) %>%
      summarize(b_u = sum(rating - b_i - u)/(n()+l))
    
    # compute predictions on validation set based on these above terms
    predicted_ratings <- validation %>% 
      left_join(b_i, by = "movieId") %>%
      left_join(b_u, by = "userId") %>%
      mutate(pred = u + b_i + b_u) %>%
      pull(pred)

    # output RMSE of these predictions
    return(RMSE(predicted_ratings, validation$rating))
  })

  # quick plot of RMSE vs lambdas
  qplot(lambdas,rmses)
  # print minimum RMSE 
  min(rmses)



```

 Final model with the best fitted lambda 
 
```{r}

lam <- lambdas[which.min(rmses)]
  
  b_i <- edx %>% 
    group_by(movieId) %>%
    summarize(b_i = sum(rating - u)/(n()+lam))
  # compute regularize user bias term
  b_u <- edx %>% 
    left_join(b_i, by="movieId") %>%
    group_by(userId) %>%
    summarize(b_u = sum(rating - b_i - u)/(n()+lam))
  
  # compute predictions on validation set based on these above terms
  predicted_ratings <- validation %>% 
    left_join(b_i, by = "movieId") %>%
    left_join(b_u, by = "userId") %>%
    mutate(pred = u + b_i + b_u) %>%
    pull(pred)
  
  # Let's find the RMSE based on the above terms
  RMSE(predicted_ratings, validation$rating)
  

```
 